The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-21 Writeup/paper_v3.tex
Distortion / Ellipticity disambiguity fixed.
Redshift distribution for volume-limited samples included

The estimated overdensity in a redshift bin is defined by comparing
the observed galaxy counts in the bin with the counts that are
predicted in that bin by one of the models in Eqs.~\eqref{E:pz1}
and~\eqref{E:pz2}:
\begin{equation}
\delta_{g,\text{1D}}=\frac{(N-N_{\text{mod}})}{N_{\text{mod}}},
\end{equation}
where
\begin{equation}
N_\text{mod} = \int_{z_\text{min}}^{z_\text{max}} p(z) \,\mathrm{d}z
\end{equation}
is determined by integrating the redshift distribution within the
limits of that redshift slice. 
%(\rachel{This $\delta$ definition is important, so it should be in an equation environment.})
Note that $\delta_{g,\text{1D}}$ is dependent on our choice of model
redshift distribution, and should have a mean value of $0$ over the
entire redshift range. \arun{weighted by the number fraction in each bin.} % If a model and the distribution agree, then $\delta_{g,\text{1D}}=0$ for that model at all the redshifts.

Our decision criterion for identifying overdense and underdense
redshift slices involves leaving a 10 per cent margin around an
overdensity of zero; i.e., if $\left| \delta_{g,\text{1D}} \right| <
0.1$, that is considered ``neutral'' (neither overdense nor
underdense on average).  We can then label each redshift slice as
either overdense, underdense, or neutral as follows: 
We label a redshift bin as overdense if at least one model gives a
value of $\delta_{g,\text{1D}} > 0.1$ while the other gives
$\delta_{g,\text{1D}} > -0.1$ 
(neutral or overdense), and vice versa for the underdense regions. We
label a redshift bin as neutral if both models give
$\delta_{g,\text{1D}}$ within the neutral region,  
\emph{or} if use of one model redshift distribution results in the
conclusion that the bin is overdense while the over leads to the
conclusion that it is underdense.
We thus identify the regions $z=0.30-0.40$, $0.65-0.75$, and
$0.80-0.85$ as overdense; $z=0.55-0.65$ and $0.75-0.80$ as underdense; and $z=0.40-0.55$ as unclassified. 
%Although, the $z=0.40-0.55$ bin appears to be underdense, for reasons mentioned in \S\ref{sub:volumelimiting}, we will not call them underdense.

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_vollimited_withcoil}
  \caption{
Upper panel: Redshift distribution of volume-limited ($M_I <  -20.8$)
\s$1$ sample with photometric redshift bins that are 0.05 wide. Two
analytical functions with best fit parameters are plotted over it. 
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) =
           N/N_{\text{mod}}$ with each functional form as the model
           for each redshift bin. \rachel{I am quite concerned that
             something has gone wrong with this fit (in the sense that
           there may be a bug).  Why is the redshift distribution
           shown in blue forced to zero at $z=0.3$?  It should go to
           zero at $z=0$, not $0.3$.  It is clearly not normalized to the same total number as the redshift histogram, either.  In terms of formatting, please
           implement the same changes that I suggested for the
           flux-limited sample redshift distribution plot.}\arun{This will change once I discuss with you.}} 
  \label{fig:redshift_vollimited}
\end{figure}

@@

\subsection{Wider bins}

We now repeat the analysis with wider redshift bins. Pretending to be ignorant of any overdensities and underdensities along the line of sight, we choose the slices in redshift to be $\left[ 0.3 - 0.5 \right]$,
$\left[ 0.5 - 0.7\right]$ and $\left[ 0.7 - 0.85\right]$.
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution_wide} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip2_Bbandevolution_wide}
 \caption{ RMS ellipticities as a function of redshift. The horizontal errorbars indicate
           the width of the redshift bin, while the vertical ones are $1\sigma$ errorbars obtained by
           bootstrapping. Points with open centers and thick errorbars correspond to the stellar-mass selected sample \s$3$
           and points with filled centers and thin errorbars correspond to the luminosity-selected samples \s$1$ and \s$2$.    
         }
 \label{fig:rms_ellip_wide}
\end{figure}

@@

Comparing Figs.~\ref{fig:rms_ellip_wide}, ~\ref{fig:rms_ellip_momentbased_wide}, ~\ref{fig:median_dvcbtt_wide} and ~\ref{fig:median_sersicn_wide} to the corresponding figures from earlier, we observe that the 
range of values they assume have become smaller and so have the vertical errorbars, due to the increase in the number of galaxies in each redshift bin. The results do not suggest that statistic values
are all consistent across all redshift bins. However, it seems to suggest that when we choose wider bins, which include overdense and underdense structures in them and thus average out,
the shape and the morphological parameters are affected lesser by the cosmic variance.

>>>>

The estimated overdensity in a redshift bin is defined by comparing
the observed galaxy counts in the bin with the counts that are
predicted in that bin by one of the models in Eqs.~\eqref{E:pz1}
and~\eqref{E:pz2}:
\begin{equation}
\delta_{g,\text{1D}}=\frac{(N-N_{\text{mod}})}{N_{\text{mod}}},
\end{equation}
where
\begin{equation}
N_\text{mod} = \int_{z_\text{min}}^{z_\text{max}} p(z) \,\mathrm{d}z
\end{equation}
is determined by integrating the redshift distribution within the
limits of that redshift slice. 
%(\rachel{This $\delta$ definition is important, so it should be in an equation environment.})
Note that $\delta_{g,\text{1D}}$ is dependent on our choice of model
redshift distribution, and should have a mean value of $0$ over the
entire redshift range when weighted by the number fraction in each bin.
%\arun{weighted by the number fraction in each bin.} 
% If a model and the distribution agree, then $\delta_{g,\text{1D}}=0$ for that model at all the redshifts.

Our decision criterion for identifying overdense and underdense
redshift slices involves leaving a 10 per cent margin around an
overdensity of zero; i.e., if $\left| \delta_{g,\text{1D}} \right| <
0.1$, that is considered ``neutral'' (neither overdense nor
underdense on average).  We can then label each redshift slice as
either overdense, underdense, or neutral as follows: 
We label a redshift bin as overdense if at least one model gives a
value of $\delta_{g,\text{1D}} > 0.1$ while the other gives
$\delta_{g,\text{1D}} > -0.1$ 
(neutral or overdense), and vice versa for the underdense regions. We
label a redshift bin as neutral if both models give
$\delta_{g,\text{1D}}$ within the neutral region,  
\emph{or} if use of one model redshift distribution results in the
conclusion that the bin is overdense while the over leads to the
conclusion that it is underdense. \arun{New:}The naive Poisson uncertainty of the counts
in each bin is much lesser compared the difference between the actual number of galaxies present
and the number predicted by the models. For this reason, errorbars have been completely ignored.
We thus identify the regions $z=0.30-0.40$, $0.65-0.75$, and
$0.80-0.85$ as overdense; $z=0.55-0.65$ and $0.75-0.80$ as underdense; and $z=0.40-0.55$ as unclassified. 
%Although, the $z=0.40-0.55$ bin appears to be underdense, for reasons mentioned in \S\ref{sub:volumelimiting}, we will not call them underdense.

@@

\begin{figure}
 \centering
  \includegraphics[width=\columnwidth]{redshift_vollimited}
  \caption{
Upper panel: Redshift distribution of volume-limited ($M_I <  -20.8$)
\s$1$ sample with photometric redshift bins that are 0.05 wide. Two
analytical functions with best fit parameters are plotted over it. 
           Lower panel: Plot of $(1+\delta_{g,\text{1D}}) =
           N/N_{\text{mod}}$ with each functional form as the model
           for each redshift bin. \rachel{I am quite concerned that
             something has gone wrong with this fit (in the sense that
           there may be a bug).  Why is the redshift distribution
           shown in blue forced to zero at $z=0.3$?  It should go to
           zero at $z=0$, not $0.3$.  It is clearly not normalized to the same total number as the redshift histogram, either.  In terms of formatting, please
           implement the same changes that I suggested for the
           flux-limited sample redshift distribution plot.}\arun{This will change once I discuss with you.}} 
  \label{fig:redshift_vollimited}
\end{figure}

@@

\subsection{Rectifying line of sight fluctuations}

\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{redshift_fluxlimited_wide}
 \caption{Plot of $\delta_{g,\text{1D}} =  N/N_{\text{mod}}-1$ with each functional form as the model
           for each redshift bin. }
 \label{fig:redshift_fluxlimited_wide}          
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{redshift_vollimited_wide}
 \caption{Plot of $\delta_{g,\text{1D}} =  N/N_{\text{mod}}-1$ with each functional form as the model
           for each redshift bin. }
 \label{fig:redshift_vollimited_wide}          
\end{figure}

\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution_wide} \
 \includegraphics[width=1.0\columnwidth]{rms_ellip1_Bbandevolution_wide}
 \caption{ RMS ellipticities as a function of redshift. The horizontal errorbars indicate
           the width of the redshift bin, while the vertical ones are $1\sigma$ errorbars obtained by
           bootstrapping. Points with open centers and thick errorbars correspond to the stellar-mass selected sample \s$3$
           and points with filled centers and thin errorbars correspond to the luminosity-selected samples \s$1$ and \s$2$.    
         }
 \label{fig:rms_ellip_wide}
\end{figure}

@@

We now repeat the analysis with wider redshift bins. Pretending to be ignorant of any overdensities and underdensities along the line of sight, we choose the slices in redshift to be $\left[ 0.3 - 0.5 \right]$,
$\left[ 0.5 - 0.7\right]$ and $\left[ 0.7 - 0.85\right]$.

Using the fits done earlier (cf. Figs.~\ref{fig:redshift_fluxlimited} and ~\ref{fig:redshift_vollimited}), we obtain the overdensities for the wider bins.
%that are shown in Figs.~\ref{fig:redshift_fluxlimited_wide} and ~\ref{fig:redshift_vollimited_wide}. 
Although the $\left[ 0.7 - 0.85\right]$ bin seems to be overdense in Fig.~\ref{fig:redshift_fluxlimited_wide}, it appears to closely follow our models when we volume-limit the sample.
See Fig.~\ref{fig:redshift_vollimited_wide}. Thus, as before, we classify $\left[ 0.7 - 0.85\right]$ bin as `neutral'.

Comparing Figs.~\ref{fig:rms_ellip_wide}, ~\ref{fig:rms_ellip_momentbased_wide}, ~\ref{fig:median_dvcbtt_wide} and ~\ref{fig:median_sersicn_wide} to the corresponding figures from earlier, we observe that the 
range of values they assume have become smaller and so have the vertical errorbars, partly due to the increase in the number of galaxies in each redshift bin. The results do not suggest that statistic values
are all consistent across all redshift bins. However, it seems to suggest that when we choose wider bins, which include overdense and underdense structures in them and thus average out,
the shape and the morphological parameters are affected lesser by the cosmic variance.
